---
title: "Uncertainty"
author: "Giuseppe Tinti Tomio"
output:
    xaringan::moon_reader:
        self_contained: yes
        nature:
            highlightStyle: github
            highlightLines: true
            countIncrementalSlides: false
            slideNumberFormat: "%current%"
---

```{r R setup, include = FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    fig.align = "left",
    fig.height = 7,
    fig.width = 10,
    fig.retina = 5,
    out.height = "100%",
    cache = TRUE
)

library(tidyverse)
library(xkcd)

COLUMN_COUNT <- 4
ROW_COUNT <- 4
TOTAL_COUNT <- COLUMN_COUNT * ROW_COUNT

EXPERIMENT_COUNT <- 100
POPULATION_SIZE <- 20
TRUE_LOCATION <- 170
TRUE_SCALE <- 10

SEED <- 42

# RED <- scales::hue_pal()(2)[1]
# BLUE <- scales::hue_pal()(2)[2]
ORANGE <- "#E69F00"
GREEN <- "#009E73"
BLUE <- "#0072B2"
BASE_SIZE <- 24

plot_stickmen <- function(
    column_count,
    row_count,
    is_selected = FALSE
) {
    data <- generate_stickmen_data(column_count, row_count, is_selected)

    plot <-
        ggplot() +
        xkcdman(aes(
            x = x,
            y = y,
            scale = draw_height,
            ratioxy = xy_ratio,
            angleofspine = spine_angle,
            anglelefthumerus = lefth_humerus_angle,
            anglerighthumerus = right_humerus_angle,
            angleleftradius = left_radius_angle,
            anglerightradius = right_radius_angle,
            angleleftleg = left_leg_angle,
            anglerightleg = right_leg_angle,
            angleofneck = neck_angle,
            color = is_selected
        ), data) +
        scale_x_continuous(expand = expansion(add = 0)) +
        scale_y_continuous(expand = expansion(add = 0)) +
        scale_color_manual(values = c(`TRUE` = ORANGE, `FALSE` = BLUE), guide = "none") +
        theme_void()

    return(plot)
}

generate_stickmen_data <- function(
    column_count,
    row_count,
    is_selected
) {
    total_count <- column_count * row_count

    x <- rep(seq(0, column_count - 1), times = row_count)
    y <- - 4 * rep(seq(0, row_count - 1), each = column_count)
    draw_height <- 1
    xy_ratio <- diff(range(x)) / diff(range(y))

    down <- - pi / 2
    tilt <- pi / 12
    left_tilted <- down + tilt
    right_tilted <- down - tilt

    data <- tibble(
        x = x,
        y = y,
        draw_height = draw_height,
        xy_ratio = xy_ratio,
        spine_angle = down,
        lefth_humerus_angle = left_tilted,
        right_humerus_angle = right_tilted,
        left_radius_angle = left_tilted,
        right_radius_angle = right_tilted,
        left_leg_angle = left_tilted,
        right_leg_angle = right_tilted,
        neck_angle = down,
        is_selected = is_selected
    )

    return(data)
}

plot_ci <- function(
    confidence_level,
    experiment_count,
    population_size,
    true_location,
    true_scale,
    seed = NULL
) {
    data <-
        confidence_level %>%
        generate_ci_data(
            experiment_count,
            population_size,
            true_location,
            true_scale,
            seed
        )

    y_breaks <- seq(true_location - 2 * true_scale, true_location + 2 * true_scale, 10)
    color_values <- c(`TRUE` = GREEN, `FALSE` = ORANGE)

    plot <-
        data %>%
        ggplot(aes(x = index, ymin = lower_ci, ymax = upper_ci, color = is_covering)) +
        geom_hline(yintercept = true_location, size = 2) +
        geom_errorbar(size = 2, width = 0) +
        scale_x_continuous(breaks = NULL, expand = expansion(add = 1)) +
        scale_y_continuous(breaks = y_breaks, labels = ~ str_glue("{.x}\ncm"), limits = range(y_breaks), expand = expansion(add = 1)) +
        scale_color_manual(values = color_values, guide = "none") +
        labs(
            x = NULL,
            y = NULL
        ) +
        theme_minimal(BASE_SIZE)

    return(plot)
}

generate_ci_data <- function(
    confidence_level,
    experiment_count,
    population_size,
    true_location,
    true_scale,
    seed = NULL
) {
    set.seed(seed)

    data <- tibble(
        index = seq(experiment_count),
        sample = replicate(experiment_count, list(rnorm(population_size, true_location, true_scale)))
    )

    data <-
        data %>%
        rowwise() %>%
        mutate(location = mean(sample), scale = sd(sample)) %>%
        ungroup()

    quantile <- 1 - (1 - confidence_level) / 2

    data <-
        data %>%
        mutate(
            lower_ci = location - qt(quantile, population_size) * scale / sqrt(population_size),
            upper_ci = location + qt(quantile, population_size) * scale / sqrt(population_size)
        ) %>%
        mutate(is_covering = (lower_ci <= true_location) & (true_location <= upper_ci))

    return(data)
}

set.seed(SEED)
```

```{css css setup}
.pull-left-70 {
    float: left;
    width: 65%;
}
.pull-right-30 {
    float: right;
    width: 35%;
}

.big {
    font-size: 25px;
}

.left {
    float : left;
}

.center {
    float: center;
}


.right {
    float : right;
}

.orange {
    color: #E69F00;
}

.green {
    color: #009E73;
}

.blue {
    color: #0072B2;
}
```

```{r reference setup, include = FALSE, cache = FALSE}
library(RefManageR)

bibliography <- ReadBib("./bibliography.bib")

BibOptions(
    bib.style = "authoryear",
    max.names = 1,
    cite.style = "authoryear",
    style = "markdown",
    hyperlink = FALSE
)

# Citet(bibliography, "shields2011bias")
```

```{r processing, include = FALSE}
```

# Uncertainty

- What is it?
- Why is it important?
- How can I measure it?
- How to deal with it?

---
class: center, middle, inverse

# A quick story

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = FALSE
)
```
]

.pull-right-30.big.center[
Height?
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = c(TRUE, rep_len(FALSE, TOTAL_COUNT - 1))
)
```
]

.pull-right-30.big.center[
.orange[Height = 171]
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = c(rep_len(FALSE, 1), TRUE, rep_len(FALSE, TOTAL_COUNT - 1 - 1))
)
```
]

.pull-right-30.big.center[
.orange[Height = 189]
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = c(rep_len(FALSE, 2), TRUE, rep_len(FALSE, TOTAL_COUNT - 2 - 1))
)
```
]

.pull-right-30.big.center[
.orange[Height = 153]
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = FALSE
)
```
]

.pull-right-30.big.center[
]

---
# What is it?

- **Certainty** is how confident you are about some guess
- **Uncertainty** is the opposite of certainty

---
# Why is it important?

> Statistic does not tell us wheter we are right. It tells us the chance of being wrong.
>
> `r Citet(bibliography, "krzywinski2013importance")`


--
- Understand uncertainty is to understand how much you know about the unkown
- Make honest guesses

--
- Job security for statistician

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = FALSE
)
```
]

.pull-right-30.big.center[
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = TRUE
)
```
]

.pull-right-30.big.center.middle[
<br>
<br>
<br>
```{r, fig.height = 8, fig.width = 6}
histogram_data <- tibble(
    height = rnorm(1000, TRUE_LOCATION, TRUE_SCALE)
)

histogram_data %>%
    ggplot(aes(x = height)) +
    geom_histogram(bins = 50, fill = ORANGE) +
    theme_void()
```
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = TRUE
)
```
]

.pull-right-30.big.center[
.orange[Bottom 1% = 140]

.orange[Top 1% = 200]
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = TRUE
)
```
]

.pull-right-30.big.center[
.orange[Mean = 170]

.orange[SD = 10]

<br>

.orange[$$\mu = \frac{\sum\limits_i x_i}{N}$$]

.orange[$$\sigma = \sqrt{\frac{\sum\limits_i (x_i - \mu)^2}{N}}$$]
]

---
# How to measure it?

- Standard deviation

$$\sigma = \sqrt{\frac{\sum\limits_i (x_i - \mu)^2}{N}}$$

--
- Median absolute deviation

$$\text{MAD} = \text{median}(|x_i - \text{median}(x)|)$$

--
- Other approaches like entropy

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = TRUE
)
```
]

.pull-right-30.big.center[
.orange[Mean = 177]

.orange[SD = 10]

<br>

.orange[$$\mu = \frac{\sum\limits_i x_i}{N}$$]

.orange[$$\sigma = \sqrt{\frac{\sum\limits_i (x_i - \mu)^2}{N}}$$]
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = sample(c(rep(TRUE, times = 4), rep(FALSE, times = TOTAL_COUNT - 4)))
)
```
]

.pull-right-30.big.center[
Pop Mean = 170

Pop SD = 10

<br>

.orange[Sample mean = 169]

.orange[Sample SD = 10]

<br>

.orange[$$\hat{\mu} = \frac{\sum\limits_i x_i}{n}$$]

.orange[$$\hat{\sigma} = \sqrt{\frac{\sum\limits_i (x_i - \hat{\mu})^2}{n - 1}}$$]
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = sample(c(rep(TRUE, times = 4), rep(FALSE, times = TOTAL_COUNT - 4)))
)
```
]

.pull-right-30.big.center[
Pop Mean = 170

Pop SD = 10

<br>

.orange[Sample mean = 163]

.orange[Sample SD = 12]

<br>

.orange[$$\hat{\mu} = \frac{\sum\limits_i x_i}{n}$$]

.orange[$$\hat{\sigma} = \sqrt{\frac{\sum\limits_i (x_i - \hat{\mu})^2}{n - 1}}$$]
]

---
# Convergence

```{r fig.height = 8, fig.width = 12}
population_data <- rnorm(1000, TRUE_LOCATION, TRUE_SCALE)

convergence_data <- tibble(
    sample_size = seq(2, 1000)
)

convergence_data <-
    convergence_data %>%
    rowwise() %>%
    mutate(mean = mean(population_data[sample(sample_size)])) %>%
    mutate(standard_deviation = sd(population_data[sample(sample_size)])) %>%
    ungroup()

y_breaks_mean <- seq(TRUE_LOCATION - 2 * TRUE_SCALE, TRUE_LOCATION + 2 * TRUE_SCALE, 10)

convergence_data %>%
    ggplot(aes(x = sample_size, y = mean)) +
    geom_hline(yintercept = TRUE_LOCATION, size = 2) +
    geom_point(color = ORANGE) +
    scale_y_continuous(breaks = y_breaks_mean, labels = ~ str_glue("{.x}\ncm"), limits = range(y_breaks_mean)) +
    theme_minimal(BASE_SIZE) +
    labs(
        x = "Sample size",
        y = "Sample mean"
    )
```

---
# Convergence

```{r fig.height = 8, fig.width = 12}
y_breaks_sd <- seq(TRUE_SCALE - 1 * TRUE_SCALE, TRUE_SCALE + 1 * TRUE_SCALE, 10)

convergence_data %>%
    ggplot(aes(x = sample_size, y = standard_deviation)) +
    geom_hline(yintercept = TRUE_SCALE, size = 2) +
    geom_point(color = ORANGE) +
    scale_y_continuous(breaks = y_breaks_sd, labels = ~ str_glue("{.x}\ncm"), limits = range(y_breaks_sd)) +
    theme_minimal(BASE_SIZE) +
    labs(
        x = "Sample size",
        y = "Sample standard deviation"
    )
```

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = sample(c(rep(TRUE, times = 4), rep(FALSE, times = TOTAL_COUNT - 4)))
)
```
]

.pull-right-30.big.center[
Pop Mean = 170

Pop SD = 10

<br>

.orange[Sample mean = 163]

.orange[Sample SD = 12]

<br>

.orange[$$\hat{\mu} = \frac{\sum\limits_i x_i}{n}$$]

.orange[$$\hat{\sigma} = \sqrt{\frac{\sum\limits_i (x_i - \hat{\mu})^2}{n - 1}}$$]
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = sample(c(rep(TRUE, times = 4), rep(FALSE, times = TOTAL_COUNT - 4)))
)
```
]

.pull-right-30.big.center[
Pop Mean = 170

Pop SD = 10

<br>

.orange[Sample mean = 163]

.orange[Sample SD = 12]

.orange[CI = [151, 176]]

<br>

.orange[$$\text{CI} = [\widehat{\mu} - \widehat{\sigma},\, \widehat{\mu} + \widehat{\sigma}]$$]
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = sample(c(rep(TRUE, times = 4), rep(FALSE, times = TOTAL_COUNT - 4)))
)
```
]

.pull-right-30.big.center[
Pop Mean = 170

Pop SD = 10

<br>

.orange[Sample mean = 181]

.orange[Sample SD = 10]

.orange[CI = [171, 191]]

<br>

.orange[$$\text{CI} = [\widehat{\mu} - \widehat{\sigma},\, \widehat{\mu} + \widehat{\sigma}]$$]
]

---
.pull-left-70[
```{r, fig.height = 8, fig.width = 6}
plot_stickmen(
    column_count = COLUMN_COUNT,
    row_count = ROW_COUNT,
    is_selected = sample(c(rep(TRUE, times = 4), rep(FALSE, times = TOTAL_COUNT - 4)))
)
```
]

.pull-right-30.big.center[
Pop Mean = 170

Pop SD = 10

<br>

.orange[Sample mean = 181]

.orange[Sample SD = 10]

.orange[CI = [161, 201]]

<br>

.orange[$$\text{CI} = [\widehat{\mu} - 2 \widehat{\sigma},\, \widehat{\mu} + 2 \widehat{\sigma}]$$]
]

---
```{r, fig.height = 8, fig.width = 12}
plot_ci(0.50, EXPERIMENT_COUNT, POPULATION_SIZE, TRUE_LOCATION, TRUE_SCALE, SEED)
```

.right.green[Confidence level `r scales::percent(0.50)`]

---
```{r, fig.height = 8, fig.width = 12}
plot_ci(0.90, EXPERIMENT_COUNT, POPULATION_SIZE, TRUE_LOCATION, TRUE_SCALE, SEED)
```

.right.green[Confidence level `r scales::percent(0.90)`]

---
# Example

This:
- `r Citet(bibliography, "shields2011bias")` report a 95% CI of [174cm, 175cm] for the height of the average Canadian man

Means:
- The average Canadian men is estimated to be between 174cm and 175cm tall
- We can be confident of this estimate, because if a lot of scientists replicated the experiment, we would expect 95% of them to be correct

---
# Reference, code and others

## Reference

```{r references, echo = FALSE, results = "asis", cache = FALSE}
PrintBibliography(bibliography)
```

## Slide source code

[github.com/GiuseppeTT/TA-presentation-module](https://github.com/GiuseppeTT/TA-presentation-module)
